Cultural Transmission Can Explain the Evolution of Cooperative Behavior

### 2020-07-15 ms/main.tex
fix beta_1 beta_3 when TA>TB

In this case $0<v<1$, and the equation system is more complex.
Therefore, we focus on local, rather than global, stability.\\
To proceed, we note that 
eq.~\ref{eq:horizontal} can give $\hat{p}'$ as a function of both $p'$ and $\tilde{p}'$,
eq.~\ref{eq:nextgen_adults_slimpify} gives $p'$ as a function of $\hat{p}$, and 
eq.~\ref{eq:nextgen_parents_simplified} gives $\tilde{p}'$ as a function of $\hat{p}$. 
Combining these equations, we find an equation for $\hat{p}'$ as a function of $\hat{p}$.
We then determine the equilibria of this equation, that is, solutions for $\hat{p}' = \hat{p}$, and analyse their local stability: an equilibrium $\hat{p}^*$ is locally stable when the derivative of $f(\hat{p})$ at the equilibrium is negative, $f'(\hat{p}^*)<0$.

@@

Note that the sign of the cubic $f(\hat{p})$ at positive (negative) infinity is equal (inverse) to the sign of $\beta_1$. 
If $T_A>T_B$, then 
\begin{equation} \label{eq:beta1}
   \beta_1 %& = [c(1-v) - b (1-\alpha v)] (T_A-T_B) \\
   < [c(1-\alpha v) - b(1-\alpha v)] (T_A-T_B) 
   = (1-\alpha v)(c-b)(T_A-T_B) < 0 ,
 \end{equation}
since $c<b$, and $1>\alpha v$.
So the sign of the cubic at positive and negative infinity is negative and positive, respectively.
First, if $\beta_3<\beta_1$ then $0<\hat{p}^*<1$ and therefore $f'(0)<0$ and $f'(1)<0$, that is, both fixations are locally stable and $\hat{p}^*$ separates the domains of attraction.
Second, if $\beta_1<\beta_3<0$ then $1<\hat{p}^*$ and therefore $f'(0)<0$ and $f'(1)>0$, that is, fixation of the defector phenotype $B$ is the only locally stable legitimate (i.e. between 0 and 1) equilibrium.
Third, if $0<\beta_3$ then $\hat{p}^*<0$ and therefore $f'(0)>0$ and $f'(1)<0$, that is, fixation of the cooperator phenotype $A$ is the only locally stable legitimate equilibrium.

@@

The following result summarizes these findings.

@@

\item \emph{Fixation of the defector phenotype $B$}, if $T_A>T_B$ and $\beta_1<\beta_3<0$, or $T_A<T_B$ and $\beta_3<0$.

@@

\item \emph{Fixation of either phenotype depending on initial frequency}, if $T_A>T_B$ and $\beta_3<\beta_1$.

>>>>

In this case $0<v<1$, and the equation system is more complex.
Therefore, we focus on local, rather than global, stability.\\
To proceed, we note that 
eq.~\ref{eq:horizontal} can give $\hat{p}'$ as a function of both $p'$ and $\tilde{p}'$,
eq.~\ref{eq:nextgen_adults_slimpify} gives $p'$ as a function of $\hat{p}$, and 
eq.~\ref{eq:nextgen_parents_simplified} gives $\tilde{p}'$ as a function of $\hat{p}$. 
Combining these equations, we find an equation for $\hat{p}'$ as a function of $\hat{p}$.
We then determine the equilibria of this equation, that is, solutions for $\hat{p}' = \hat{p}$, and analyse their local stability: an equilibrium $\hat{p}^*$ is locally stable when the derivative of $f(\hat{p})$ at the equilibrium is negative, $f'(\hat{p}^*)<0$. %TODO show this is necessary and sufficient

@@

Note that the sign of the cubic $f(\hat{p})$ at positive (negative) infinity is equal (inverse) to the sign of $\beta_1$. 
If $T_A>T_B$, then 
\begin{equation} \label{eq:beta1}
   \beta_1 %& = [c(1-v) - b (1-\alpha v)] (T_A-T_B) \\
   < [c(1-\alpha v) - b(1-\alpha v)] (T_A-T_B) 
   = (1-\alpha v)(c-b)(T_A-T_B) < 0 ,
 \end{equation}
since $c<b$, and $1>\alpha v$.
So the sign of the cubic at positive and negative infinity is negative and positive, respectively.
First, if $\beta_3<\beta_1$ then 
$1<\hat{p}^*$ and therefore $f'(0)<0$ and $f'(1)>0$, that is, fixation of the defector phenotype $B$ is the only locally stable legitimate (i.e. between 0 and 1) equilibrium.
Second, if $\beta_1<\beta_3<0$ then 
$0<\hat{p}^*<1$ and therefore $f'(0)<0$ and $f'(1)<0$, that is, both fixations are locally stable and $\hat{p}^*$ separates the domains of attraction.
Third, if $0<\beta_3$ then 
$\hat{p}^*<0$ and therefore $f'(0)>0$ and $f'(1)<0$, that is, fixation of the cooperator phenotype $A$ is the only locally stable legitimate equilibrium.

@@

The following result summaries these findings.

@@

\item \emph{Fixation of the defector phenotype $B$}, if $T_A>T_B$ and $\beta_3<\beta_1$, or $T_A<T_B$ and $\beta_3<0$.

@@

\item \emph{Fixation of either phenotype depending on initial frequency}, if $T_A>T_B$ and $\beta_1<\beta_3<0$.
